\documentclass[10pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{hyperref}

\title{mcp-stdio-guard: Protocol Hygiene Evaluation for MCP Stdio Servers}
\author{}
\date{}

\begin{document}
\maketitle

\begin{abstract}
Model Context Protocol (MCP) stdio servers use stdout as a JSON-RPC protocol
channel. This paper studies failure modes caused by stdout pollution,
JSON-RPC framing errors, incomplete MCP lifecycle behavior, process crashes,
tool metadata defects, capability drift, and strict protocol edge cases. The
artifact evaluates a pinned \texttt{mcp-stdio-guard} release against synthetic
fixtures and records all reported results as reproducible machine-readable
outputs.
\end{abstract}

\section{Introduction}

MCP stdio transports reserve stdout for protocol messages and use stderr for
diagnostics \cite{mcpstdio}. This boundary is easy to violate in practical
server implementations: startup banners, logging defaults, progress bars,
debugging \texttt{print} statements, malformed frames, and process crashes can
all surface as confusing client failures. The goal of this work is to make that
boundary observable and reproducible for server authors and registry operators.

This paper evaluates \texttt{mcp-stdio-guard}, a command-line validation tool
for MCP stdio servers \cite{mcpstdioguard}. The accompanying artifact pins a
tool release, runs synthetic MCP stdio fixtures through multiple guard profiles,
and generates the result tables from raw command output.

\section{Background}

JSON-RPC 2.0 defines request, response, notification, id, and error-object
semantics \cite{jsonrpc20}. MCP builds on JSON-RPC and defines an initialize
lifecycle where clients and servers negotiate protocol version and capabilities
before normal operation \cite{mcplifecycle}. In the stdio transport, stdout is
therefore not a general-purpose logging stream; it is the protocol channel.

\section{Problem Statement}

The research question for this artifact is:

\begin{quote}
Can a lightweight external guard detect common MCP stdio protocol hygiene
failures without modifying the target server?
\end{quote}

The artifact focuses on observable behavior at the process boundary. It does not
attempt to prove semantic correctness of arbitrary tool implementations.

\section{Failure Taxonomy}

The evaluation groups injected faults into the following categories:

\begin{itemize}
  \item stdout transport pollution, including startup and late diagnostic output;
  \item invalid JSON-RPC framing, including malformed JSON, LSP-style
        \texttt{Content-Length} framing, and response id mismatch;
  \item MCP lifecycle failures, including missing initialize responses and
        crashes after the initialized notification;
  \item registry-facing metadata failures, including duplicate tool names and
        invalid tool input schemas;
  \item repeat-run instability, including capability and tool-list drift;
  \item strict adversarial behavior, including probes that expect structured
        JSON-RPC errors or no response for notifications.
\end{itemize}

\section{System Design}

\texttt{mcp-stdio-guard} runs a target server as a child process, performs the
MCP initialize lifecycle, optionally sends post-initialize operations, validates
stdout as protocol-only JSON-RPC output, records process health, and emits a
machine-readable JSON result. Profiles select different depths of checking:
smoke, registry, CI, and strict. Registry and strict profiles exercise repeat
runs; CI and strict can include static stdout-write scanning; strict adds
adversarial protocol probes.

\section{Evaluation Methodology}

The artifact evaluates \texttt{mcp-stdio-guard@1.0.0}. The command used for the
official snapshot is:

\begin{verbatim}
MCP_STDIO_GUARD_VERSION=mcp-stdio-guard@1.0.0 npm run evaluate
\end{verbatim}

The evaluation script runs synthetic server fixtures under relevant guard
profiles and writes raw JSON, CSV, and Markdown summaries. The official baseline
snapshot is preserved under
\texttt{artifact/results/v1.0.0-paper-baseline/}. Tables in this paper are
generated from that snapshot.

\section{Results}

Table~\ref{tab:detection-matrix} summarizes which injected faults were detected
and which issue codes appeared in the official baseline. Table~\ref{tab:runtime-overhead}
summarizes observed wall-clock runtime by profile. These tables are generated
from actual evaluation output.

\input{tables/detection-matrix}

\input{tables/runtime-overhead}

\section{Discussion}

The synthetic matrix exercises protocol-boundary failures that are small enough
to reproduce but representative of practical MCP stdio mistakes. The results
separate expected pass cases, such as stderr-only diagnostics, from protocol
failures, runtime failures, metadata failures, and repeat-run drift. This
separation is important for registry display because an install/runtime issue
should not be presented as the same class of failure as stdout protocol
corruption or JSON-RPC conformance.

\section{Threats to Validity}

The current artifact uses synthetic fixtures rather than a broad real-world MCP
server corpus. Runtime measurements include process startup and \texttt{npx}
invocation overhead, so they are best interpreted as end-to-end evaluation
costs, not as isolated parser overhead. The fault taxonomy is intentionally
focused on stdio transport and lifecycle behavior; it does not evaluate semantic
tool correctness, authorization behavior, or data-source safety. Results are
also tied to the pinned tool version and should be regenerated when the guard or
MCP specification changes.

\section{Related Work}

This work sits near protocol conformance testing, language-server transport
validation, CLI integration testing, and supply-chain registry quality signals.
Unlike in-process unit tests, this artifact evaluates the server at the stdio
process boundary. Unlike static-only checks, it observes actual runtime output
and lifecycle behavior.

\section{Conclusion}

The artifact provides a reproducible baseline for evaluating MCP stdio protocol
hygiene. It preserves raw command output, generated CSV summaries, and LaTeX
tables derived from a pinned \texttt{mcp-stdio-guard} release. Future work should
extend the synthetic matrix with a curated real-server corpus and compare the
guard's findings against independent maintainer diagnoses.

\bibliographystyle{plain}
\bibliography{references}

\end{document}
